---
title: "Homework 2"
author: "Jakey Lebwohl"
date: "2023-07-08"
output: github_document
---

Always load tidyverse.

```{r}
library(tidyverse)
```

## Problem 1

Import the NYC transit data and retain relevant variables. Also clean it a little.

```{r}
nyc_subways =
  read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data (1).csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, (route1:route11), entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

This dataset contains data for line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance for subway stations in NYC. So far I have cleaned the names of the variables, retained only certain variables, and changed the values of the variable 'entry' to be logical variables instead of characters. The data set is `r nrow(nyc_subways)` rows long and `r ncol(nyc_subways)` columns wide. These data are not tidy: routes served should not be broken up into 11 variables (route1, route2, etc.): it should be broken up into route number and route name.

Keeping only distinct values of the dataset:

```{r}
nyc_subways_distinct = 
  nyc_subways %>% 
  distinct(line, station_name, .keep_all = TRUE)
```

Using this new dataframe, we can find that there are `r nrow(nyc_subways_distinct)` stations.

The number of stations which are ADA compliant is `r sum(nyc_subways_distinct$ada == TRUE)`, which can be found using the code `sum(nyc_subways_distinct$ada == TRUE)`. 

Filtering only the stations without vending:

```{r}
nyc_subways_novending = 
  nyc_subways %>% 
  filter(vending == "NO")
```

Using `sum(nyc_subways_novending$entry == TRUE)` / `nrow(nyc_subways_novending)`, we find that the proportion of station entrances / exits without vending which allow entrance is `r sum(nyc_subways_novending$entry == TRUE)` / `r nrow(nyc_subways_novending)` = `r sum(nyc_subways_novending$entry == TRUE) /  nrow(nyc_subways_novending)`.

Reformating + tidying data for routes. First we need to convert some routes from integer values to strings:

```{r}
nyc_subways_tidy = 
  nyc_subways_distinct %>% 
  transform(route8 = as.character(route8)) %>% 
  transform(route9 = as.character(route9)) %>% 
  transform(route10 = as.character(route10)) %>% 
  transform(route11 = as.character(route11)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

Filtering distinct stations which serve the A train.

```{r}
nyc_subways_atrain = 
  nyc_subways_tidy %>% 
  filter(route_name == "A") %>% 
  distinct(line, station_name, .keep_all = TRUE)
```

There are `r nrow(nyc_subways_atrain)` distinct stations which serve the A train. Of these `r sum(nyc_subways_atrain$ada)` are ADA compliant.

## Problem 2

I got bored

## Problem 3

Import and prep data

```{r}
pols_month = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  mutate(month = snakecase::to_snake_case(month.abb[as.numeric(month)])) %>% 
  mutate(year = as.numeric(year)) %>% 
  select(!c(day, prez_gop, prez_dem)) %>% 
  janitor::clean_names()

snp_month =
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year <= 15, year + 2000, year + 1900)) %>% 
  arrange(year, month) %>% 
  mutate(month = snakecase::to_snake_case(month.abb[as.numeric(month)])) %>% 
  select(!day) %>% 
  relocate(year, month) %>% 
  janitor::clean_names()


unemployment_month =
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  )

full_df = 
  left_join(pols_month, snp_month, by = c("year", "month")) %>% 
  left_join(unemployment_month, by = c("year", "month"))
  

```
